### 2023-02-14 abstract-intro/paper.tex
edited paper.tex

\documentclass[conference]{IEEEtran}
%\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}

    \title{Low cost IOT based cloud solution for Intelligence Integrated Black fly rearing.}

    \author{

        \IEEEauthorblockN{1\textsuperscript{st} Shawal Mbalire}
        \IEEEauthorblockA{\textit{School of Engineering,CEDAT} \\
        \textit{Makerere University}\\
        Kampala, Uganda \\
        mshawal49@icloud.com}

        \and
        \IEEEauthorblockN{2\textsuperscript{nd} Mary Patience Namugwanya}
        \IEEEauthorblockA{\textit{School of Engineering,CEDAT} \\
        \textit{Makerere University}\\
        Kampala, Uganda \\
        nmarypatience@gmail.com}

        \and
        \IEEEauthorblockN{3\textsuperscript{rd} Bradley Alfred Mpeera Tuwmine}
        \IEEEauthorblockA{\textit{School of Engineering,CEDAT} \\
        \textit{Makerere University}\\
        Kampala, Uganda \\
        bradleytumwine82@gmail.com}
    }

    \maketitle

    \begin{abstract}
        This paper aims to increase the production of black fly larvae for farmers to scale up by automating the monotonous and monitoring intesive tasks.
    \end{abstract}

    \begin{IEEEkeywords}
        black fly soldier fly larvea,iot
    \end{IEEEkeywords}

    \section{Introduction}

\end{document}